---
title: "Data Import"
author: "Thomas Laurent"
date: "2017年6月26日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Loading package
library(tidyverse)
library(feather)
```

#Read csv
read_ functions are much more faster and generate tibbles.

##Inline file
```{r}
read_csv("a,b,c
1,2,3
4,5,6")
```

##Skipping lines

```{r}
read_csv("The first line of metadata
  The second line of metadata
  x,y,z
  1,2,3", skip = 2)
```

```{r}
#\n adding a new line
read_csv("1,2,3\n4,5,6", col_names = FALSE)
```

```{r}
#adding colnames to csv
read_csv("1,2,3\n4,5,6", col_names = c("x", "y", "z"))
```

#Parsing a vector

```{r}
str(parse_logical(c("TRUE", "FALSE", "NA")))
str(parse_integer(c("1", "2", "3")))
str(parse_date(c("2010-01-01", "1979-10-14")))
```

##Locale

```{r}
parse_double("1,23", locale = locale(decimal_mark = ","))

parse_number("123.456.789", locale = locale(grouping_mark = "."))
```

###parse_number

Useful for extracting number in text

```{r}
parse_number("$100")
#> [1] 100
parse_number("20%")
#> [1] 20
parse_number("It cost $123.45")
```

###parsing character with special alphabets

```{r}
x1 <- "El Ni\xf1o was particularly bad this year"
x2 <- "\x82\xb1\x82\xf1\x82\xc9\x82\xbf\x82\xcd"

parse_character(x1, locale = locale(encoding = "Latin1"))
#> [1] "El Niño was particularly bad this year"
parse_character(x2, locale = locale(encoding = "Shift-JIS"))
#> [1] "こんにちは"
```

There is a function to guess which encoding would be more suitable.

```{r}
guess_encoding(charToRaw(x1))

guess_encoding(charToRaw(x2))
```

###Parsing factors

```{r}
fruit <- c("apple", "banana")
parse_factor(c("apple", "banana"), levels = fruit)
```

###Date and datetime

```{r}
parse_datetime("2010-10-01T2010")
#> [1] "2010-10-01 20:10:00 UTC"
# If time is omitted, it will be set to midnight
parse_datetime("20101010")
#> [1] "2010-10-10 UTC"
parse_date("2010-10-01")
```

For time use the package "hms"

```{r}
library(hms)
parse_time("01:10 am")
#> 01:10:00
parse_time("20:10:01")
```

Time parsing with formats
```{r}
parse_date("01/02/15", "%m/%d/%y")
#> [1] "2015-01-02"
parse_date("01/02/15", "%d/%m/%y")
#> [1] "2015-02-01"
parse_date("01/02/15", "%y/%m/%d")
#> [1] "2001-02-15"

#Using french locale
parse_date("1 janvier 2015", "%d %B %Y", locale = locale("fr"))
```

#Importing file

Guessing the type of the column
```{r}
guess_parser("2010-10-01")
```


```{r}
challenge <- read_csv(readr_example("challenge.csv"))

problems(challenge)
```

Specifying the column types when there are problems.
```{r}
challenge <- read_csv(
  readr_example("challenge.csv"), 
  col_types = cols(
    x = col_double(),
    y = col_date()
  )
)
```

Select the number of rows to determine the type of data 
```{r}
challenge2 <- read_csv(readr_example("challenge.csv"), guess_max = 1001)
```

Read data as a character columns by default
```{r}
challenge2 <- read_csv(readr_example("challenge.csv"), 
  col_types = cols(.default = col_character())
)
```

#Conversion of the data type in a tibble

```{r}
df <- tribble(
  ~x,  ~y,
  "1", "1.21",
  "2", "2.32",
  "3", "4.56"
)

type_convert(df)
```

For debugging, use read_file()

```{r}
#Reading the data as a single line
read_file(readr_example("challenge.csv"))
```

#Writing to a file
* Saving dates in ISO format
* Encoding in UTF8
```{r}
#CSV (data type is lost)
write_csv(challenge, "challenge.csv")

#Keeping data type
write_rds(challenge, "challenge.rds")

#Export the data in feather format (compatible with other programming languages)
write_feather(challenge, "challenge.feather")
```

