---
title: "Data Visualization"
output: 
  html_document:
    number_sections: true
    toc: true
---


<style type="text/css">

body{ /* Normal  */
      font-size: 12px;
  }
td {  /* Table  */
  font-size: 8px;
}
h1.title {
  font-size: 38px;
  color: DarkRed;
}
h1 { /* Header 1 */
  font-size: 28px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 22px;
  color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE,message=FALSE}
#Loading relevant packages
library(tidyverse)
library(ggplot2)
library(magrittr)
library(dplyr)
```

#Creation of simple scatter plot

To create a simple scatter plot using R, you can use ggplot() function with the dataset name as argument and geom_point() geom function with aesthetic (aes()) taking the argument of data to be plotted on each axis.
```{r}
#Simple graph
dim(mpg)
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = class, y = drv))

```

##Aesthetics

Aesthetics aes() function allows to modify the appearance of the geoms (point in the example below).
* color: change the color of geom
* size: affect the size of geom
* alpha: affect the level of transparency
* shape: change the type of shape (for certain type of geom)
* linetype: change the type of the line (with geom_smooth geom in the example below)

```{r}
#Using aesthetic to change point appearance (categorical)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy,color=class))

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy,size=class))

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy,alpha=class))

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy,shape=class))

#Different line by cat variables using linetype in aesthetic
ggplot(data = mpg,aes(x = displ, y = hwy, linetype = drv)) + 
  geom_smooth()
```

These parameters can be mapped to categorical or continuous variables.
```{r}
#Using aesthetic to change point appearance (continuous)
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy,color=cty))

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy,size=cty))

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy,alpha=cty))
```


#Using facets

Facets can be used in conjunction with ggplot to plot separately the data according to catagorical variables.

##One variable
* facet_wrap: display the different graphs in rows 
* facet_grid: display the different graphs in rows or in coloumn
```{r}
#Two rows
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  #nrow option allows to control the number of rows in the output
  facet_wrap(~ class, nrow = 2)

#One row
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(. ~ cyl)

```

##Two variables

Facets can be used by combining two different categorical variables in facet_grid() function.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(drv ~ cyl)
```

#Using different geoms

* geom_point(): scatter plot
* geom_smooth(): fitted curve
```{r}
# Scatter plot
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

# Smoothing
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy))

```

The following options can be used to plot the data for different groups when using geom_smooth():
* color: plot one curve by group (different colors)
* group: plot one curve by group

```{r}
#Plotting curves by group variable
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
              
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, group = drv))
    
ggplot(data = mpg) +
  geom_smooth(
    mapping = aes(x = displ, y = hwy, color = drv),
    show.legend = FALSE
  )
```

#Different geoms in a same plot

It is possible to use different combinations of geoms and display it together.
```{r}
#Combination of point and smooth geoms
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

#Global mapping/Local mapping

Global mapping allocates parameters to all geoms but local mapping (*i.e* declared in geom function) overrides global mapping. It can be used to plot different data within a same plot.

```{r}
#Global mapping
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
#Local mapping
  geom_point(mapping = aes(color = class)) + 
  geom_smooth()

#Subset in local mapping
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth(data = filter(mpg, class == "subcompact"), se = FALSE)
```

#Histogram

Histogram can be plotted using histogram geoms (geom_histogram() function).
```{r}
ggplot(data = mpg, mapping = aes(x = displ)) + 
  geom_histogram()
```

#Bar graph with calculated values

Another option to plot histograms is to create a dataset with respective frequencies by group and then use the geom_bar() geom function.
```{r}
#Data with frequency by a define categorical variable (tibble data frame)
demo <- tribble(
  ~cut,         ~freq,
  "Fair",       1610,
  "Good",       4906,
  "Very Good",  12082,
  "Premium",    13791,
  "Ideal",      21551
)

#Plotting the frequencies by group
ggplot(data = demo) +
  geom_bar(mapping = aes(x = cut, y = freq), stat = "identity")
```

Overall percentage using bar chart can be displayed by specifiying "y=..prop..".
```{r}
#Plotting the percentage by group (overall percentage)
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop.., group = 1))
```

#Summary statistics

Summary statistics (quantile, etc...) displayed can be specified using stat_summary() function in "fun." arguments.
```{r}
#Summary
ggplot(data = diamonds) + 
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  )
```

#Position adjustment

##Display group using colors with geom_bar
* colour: color the contour of the geom (by group)
* fill: fill the geom with the specified color (by group)
```{r}
#Adjusting how to display groups using colors
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, colour = cut))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = cut))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity))

ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) + 
  geom_bar(alpha = 1/5, position = "identity")
ggplot(data = diamonds, mapping = aes(x = cut, colour = clarity)) + 
  geom_bar(fill = NA, position = "identity")
```

##Control the position

###Geom_bar

* position="fill": bar are stacked by group variable
* position="dodge": bar are separated horizontally by group
```{r}
#Bar same size
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "fill")

#Bar horizontally splitted
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge")

```

#Geom_point

Many points may overlap on the graph and thus to have a better visualization of the distribution, position="jitter" instruction (as indicated below), adds a small noise to the data to allow to distinguish overlapping points.
```{r}

#Adding noise to avoid data point rounding
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), position = "jitter")
```

#Coordinates

Reversing x and y-axis or to display the data using polar coordinated using coord_flip() and coord_polar() functions, respectively. 
```{r}
#Change orientation of axis
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot() +
  coord_flip()

#Polar coordinates
bar <- ggplot(data = diamonds) + 
  geom_bar(
    mapping = aes(x = cut, fill = cut), 
    show.legend = FALSE,
    width = 1
  ) + 
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)

bar + coord_flip()
bar + coord_polar(theta="x") + labs(title = "Cylinders")
```

#Distribution

Density distribution can be plotted by setting y=..density.. into the aesthetic parameters.
```{r}
#Setting density instead of count for freqpoly
ggplot(data = diamonds, mapping = aes(x = price, y = ..density..)) + 
  geom_freqpoly(mapping = aes(colour = cut), binwidth = 500)
```

Boxplot can be reorder according to descriptive statistic values (median in the example below).
```{r}
#Reordering categorical variable order according to the median (FUN argument)
ggplot(data = mpg) +
  geom_boxplot(mapping = aes(x = reorder(class, hwy, FUN = median), y = hwy))
```

To display frequencies according to 2 categorical variables, two geoms are available:
* geom_tile(): data has to be processed beforehand using count function to calculate the frequencies in each group (color x cut in the example below)
* geom_bin2d()

```{r}
#Visualising the count for 2 categorical variables
diamonds %>% 
  count(color, cut) %>%  
  ggplot(mapping = aes(x = color, y = cut)) +
    geom_tile(mapping = aes(fill = n))
```

```{r}
#Creating 2d bins
ggplot(data = diamonds) +
  geom_bin2d(mapping = aes(x = carat, y = price))
```

