---
title: "Data Transformation"
output: 
  html_document:
    number_sections: true
    toc: true
---

<style type="text/css">

body{ /* Normal  */
      font-size: 12px;
  }
td {  /* Table  */
  font-size: 8px;
}
h1.title {
  font-size: 38px;
  color: DarkRed;
}
h1 { /* Header 1 */
  font-size: 28px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 22px;
  color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,comment="",message=FALSE,warning=FALSE)
```

```{r,message=FALSE,echo=FALSE}
#Package
  library(magrittr)
  library(dplyr)
  library(ggplot2)
  library(tidyr)
  library(tibble)
  library(nycflights13)
```

#Data transformation

* filter(): selecting observations based on criterion (equivalent to if...then...output in SAS)
* select(): selecting output variables (equivalent to "keep" in SAS)
* mutate(): create new variables with functions of existing variables
* summarise(): Collapse many values down to a single summary (equivalent to "PROC MEANS" in SAS)
* arrange(): sort variables by increasing or decreasing order (equivalent to "PROC SORT" in SAS)
* group_by(): perform the above instructions by variable(s) (equivalent to "by" in SAS)

Input and output of these functions are data frames or tibbles.


##Filtering basic

* filter() function take as argument the data.frame and the variable on which the condition is applied.
* near() function is useful when comparing two numbers resulting in an approximation (approximation of the root in the example below).
```{r}
filter(flights, month == 1, day == 1)
#Print and store
(dec25 <- filter(flights, month == 12, day == 25))
#Comparison floating point numbers
near(sqrt(2)^2,2)
#Filtering between two values
flights %>% filter(between(day,1,3))
```

###Missing values and list

To select missing values, is.na() function can be used within filter() function. Also, filtering using a list (vector) can be performed using the operator "%in%".
```{r}
x=NA
is.na(x)

#Display NA values
df <- tibble(x = c(1, NA, 3))
filter(df, x > 1)
#> # A tibble: 1 × 1
#>       x
#>   <dbl>
#> 1     3
filter(df, is.na(x) | x > 1)

filter(flights,month %in% c(7,8,9))

filter(flights,is.na(dep_time))
```

##Arrange (sort)

To sort data by variables, use the arrange() function. The dataset is sorted by the order of the listed variables. 
To sort in the decreasing order, use desc() function within arrange, with the variable name as argument.
```{r}
arrange(flights, year, month, day)

arrange(flights, desc(year))

##NA values are sorted at the end##

arrange(flights, is.na(arr_delay))
```

##Select

* To keep specified variables, or to reorder the order of the columns of the dataset, use the select() function. 
* The operator ":" allows to select a range of columns and "-" operator allows to delete columns.
* everything() allows to include other variables not explicitely mention in select() function
```{r}
#Selection column between two variables
select(flights, year:day)

#Opposite selection (Removing variables)
select(flights, -(year:day))

#Renaming variable (new_name=old_name)
rename(flights, tail_num = tailnum)

#Change the order of the columns
select(flights, time_hour, air_time, everything())
```

##Mutate

To create new variables, use the mutate() function which takes the data.frame and the new variable definition formula as argument. 
```{r}
#Mutate
flights_sml <- select(flights, 
  year:day, 
  ##ends_with allows to select variable whose name ends with the argument declared i.e "delay" in this example 
  ends_with("delay"), 
  distance, 
  air_time
)
mutate(flights_sml,
  gain = arr_delay - dep_delay,
  speed = distance / air_time * 60
)

```

Transmute is an alternative to mutate and performs exactly as mutate() except that the old variable is dropped from the dataframe.
```{r}
#Transmute (keep only the new variables)
transmute(flights,
  gain = arr_delay - dep_delay,
  hours = air_time / 60,
  gain_per_hour = gain / hours
)

#Divide the time into minute and hour using transmutate

transmute(flights,
  dep_time,
  hour = dep_time %/% 100,
  minute = dep_time %% 100
)

transmute(flights,
  dep_time,
  hour = dep_time%/%3600,
  minute = (dep_time-3600*hour)%/%60
)
```


##Lag function

* lag() function is similar to lag function in SAS. It takes a vector as argument and shift the data by one increment.
* lead() function performs the opposite and replace the value by the next value.
* min_rank() outputs the rank of each data (in the increasing order).
```{r}
#Lag et lead (fonction inverse)
(x <- 1:10)
lag(x)
#>  [1] NA  1  2  3  4  5  6  7  8  9
lead(x)

y <- c(1, 2,4, NA, 3, 5)
min_rank(y)
```


##Summarise

summarise() function allows to calculate overall descriptive summary,
```{r}
#Overall summary
summarise(flights, delay = mean(dep_delay, na.rm = TRUE))
```

To descriptively summarize the dataset by group use group_by() function. group_by() function takes the dataset name and the name of the grouping variable as argument.
```{r}
#Summary by group using group_by without pipeline
  by_dest <- group_by(flights, dest)
  delay <- summarise(by_dest,
    count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE)
  )
  delay
```

```{r,eval=FALSE}
#Summary by group using group_by with pipeline
  delays <- flights %>% 
    group_by(dest) %>% 
    summarise(
      count = n(),
      dist = mean(distance, na.rm = TRUE),
      delay = mean(arr_delay, na.rm = TRUE)
    )
  delays
```

You can explicitely mention to remove missing values. If missing values are present, NA will be output.
```{r}
#Removing missing values for aggregating functions (argument na.rm)
flights %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay, na.rm = TRUE))
```

Many function can be used to calculate descriptive summary (n,sd, min, quantile, median, max).
```{r}
#Other summarizing functions (sd, min, p25, median, p75, max)

flights %>% 
  group_by(dest) %>% 
  summarise(
    #n
    n=n(),
    #standard deviation
    delay_sd = sd(dep_delay,na.rm=TRUE),
    #minimum
    min=min(dep_delay,na.rm=TRUE),
    #1st quartile
    p25=quantile(dep_delay,0.25,na.rm=TRUE),
    #Median
    median=median(dep_delay,na.rm=TRUE),
    #3rd quartile
    p75=quantile(dep_delay,0.75,na.rm=TRUE),
    #Max
    max=max(dep_delay,na.rm=TRUE)) %>% 
  filter(n>25)

```

It is possible to perform the sum of a variable by a specified variable is the count() function. The first variable is the group variable and the second is the variable to be summed
```{r}
#Count data (summing the variable "distance"") by "tailnum" variable
flights %>% 
  count(tailnum, wt = distance)
```

It is possible to calculate the frequency for each variable category using the count() function.

```{r}
#Counting the frequency of dest variable
flights %>% 
  count(dest)
```

To perform the calculation of the number or the proportion of observations according to one criteria, it is possible to use sum() and mean() functions, respectively. 
```{r}
#Counting the number of observations according to one condition: dep_time < 500
flights %>% 
  group_by(year, month, day) %>% 
  summarise(n_early = sum(dep_time < 500))

#Counting the proportion of observations according to another condition arr_delay > 60
flights %>% 
  group_by(year, month, day) %>% 
  summarise(hour_perc = mean(arr_delay > 60))
```

If summarise is used in a sequential way, at the following step, data are summarized at the next above level.
```{r}
#Summary at the lowest hierarchy
daily <- group_by(flights, year, month, day)
(per_day   <- summarise(daily, flights = n()))

#Summary at the above level
(per_month <- summarise(per_day, flights = sum(flights)))
```

#Ungrouping the data

To remove to the grouping settings, use ungroup() function.
```{r}
daily <- group_by(flights, year)
#Ungrouping the data to perform overall summary for example
daily %>% 
  ungroup() %>% 
  summarise(flights = n()) 
```

#Group mutate (and filters)

It is possible to use mutate() and filter() functions downstream group_by(). 
```{r}
#Using filter function to filter by group
flights_sml %>% 
  group_by(year, month, day) %>%
  filter(rank(desc(arr_delay)) < 10)

#Filter group by size
popular_dests <- flights %>% 
  group_by(dest) %>% 
  filter(n() > 365)
popular_dests

```

```{r}

#Calculate a proportion and restrict the number of variables in the data frame
#Notice that popular_dests is a grouped data.frame
popular_dests %>% 
  filter(arr_delay > 0) %>% 
  mutate(prop_delay = arr_delay / sum(arr_delay)) %>% 
  select(year:day, dest, arr_delay, prop_delay)
```


#Using pipes

##Example of using intermediate objects

This is a way to proceed to several steps using intermediate objects. However, the code tends to be lengthy and the incrementation of each object has to be carefully done.
```{r}
#Overwriting the data
  foo_foo <- as.data.frame(c(1,2,3))
  foo_foo_1 <- cbind(foo_foo,c(2,3,4))
  foo_foo_2 <- cbind(foo_foo_1,c(3,4,5))
  foo_foo_3 <- cbind(foo_foo_2,c(4,5,6))
  colnames(foo_foo_3)=c("a","b","c","d")
  
```

##Overwriting the data

Another possibility is to overwrite the object at each step.
```{r}
#Overwriting the data
  foo_foo <- data.frame(c(1,2,3))
  foo_foo <- cbind(foo_foo,c(2,3,4))
  foo_foo <- cbind(foo_foo,c(3,4,5))
  foo_foo <- cbind(foo_foo,c(4,5,6))
  colnames(foo_foo)=c("a","b","c","d")
```

##Function composition

This method does not require any intermediate object but it may be difficult to read since it should be read from the last instruction.
```{r}
#Method using intermediate objects
  foo_foo=as.data.frame(foo_foo)
  foo_foo_1=select(foo_foo,b)
  foo_foo_2=summarise(foo_foo_1,mean=mean(b))
  
#Method using function composition (encapsulation)
  summarise(select(as.data.frame(foo_foo),b),mean=mean(b))
```

##Using pipes

Pipe are called using the symbol %>% (shortcut: Ctrl+Shift+M) and create an intermediate object "." which is discarded at last.
```{r}
#Using pipes
  foo_foo %>% 
    as.data.frame() %>% 
    select(b) %>%
    summarise(mean=mean(b))
    
#How does pipe work?
mypipe=function(.){
  .=as.data.frame(.)
  .=select(.,b)
  summarise(.,mean=mean(b))
  
}

mypipe(foo_foo)
```

##Caution when using pipe
* Pipes do not work with assign() and environment() functions.
* Better to limit the number of operations
* Not adapted to multiple inputs

#Other operators than pipes

##The %$% operator

Useful for splitting variables from a dataframe
```{r}
#Splitting variables from a dataframe
mtcars %$%
  cor(disp, mpg)
```

##The %<>% operator

Useful for storing the output into the original dataframe

```{r}
mtcars %<>% transform(cyl = cyl * 2)
```

