---
title: "Date and time"
author: "Thomas Laurent"
date: "2017年7月1日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Loading packages
library(tidyverse)
##Package lubridate to work with date and time
library(lubridate)
library(nycflights13)
```

#Creating Date/Time

* A date. Tibbles print this as <date>.

* A time within a day. Tibbles print this as <time>.

* A date-time is a date plus a time: it uniquely identifies an instant in time (typically to the nearest second). Tibbles print this as <dttm>. Elsewhere in R these are called POSIXct, but I don’t think that’s a very useful name.

Present time and date
```{r}
today()

now()
```

3 ways to create date/time

* From a string.
* From individual date-time components.
* From an existing date/time object.

##From strings

###Date
```{r}
ymd("2017-01-31")
#> [1] "2017-01-31"
mdy("January 31st, 2017")
#> [1] "2017-01-31"
dmy("31-Jan-2017")
#> [1] "2017-01-31"
ymd(20170131)
```

###Date time

```{r}
ymd_hms("2017-01-31 20:11:59")
```

##From individual components

```{r}
flights %>% 
  select(year, month, day, hour, minute) %>% 
  mutate(departure = make_datetime(year, month, day, hour, minute))
```

```{r}
#Function to convert time in hour and minute
make_datetime_100 <- function(year, month, day, time) {
  make_datetime(year, month, day, time %/% 100, time %% 100)
}

#creating date time
flights_dt <- flights %>% 
  filter(!is.na(dep_time), !is.na(arr_time)) %>% 
  mutate(
    dep_time = make_datetime_100(year, month, day, dep_time),
    arr_time = make_datetime_100(year, month, day, arr_time),
    sched_dep_time = make_datetime_100(year, month, day, sched_dep_time),
    sched_arr_time = make_datetime_100(year, month, day, sched_arr_time)
  ) %>% 
  select(origin, dest, ends_with("delay"), ends_with("time"))

#visualizing the distribution
flights_dt %>% 
  ggplot(aes(dep_time)) + 
  geom_freqpoly(binwidth = 86400) # 86400 seconds = 1 day
```

##From other types

Using as_datetime() and as_date
If a numeric value is used as an argument, reference date is 1970-01-01
```{r}
as_datetime(today())
#> [1] "2017-05-04 UTC"
as_date(now())
```


#Date components

You can extract the components of date as follows
```{r}
datetime <- ymd_hms("2016-07-08 12:34:56")

year(datetime)
#> [1] 2016
month(datetime)
#> [1] 7
mday(datetime)
#> [1] 8

#Day of the year
yday(datetime)
#> [1] 190

#Day of the week
wday(datetime)

#Displaying the name of the month using label
month(datetime, label = TRUE)

#Displaying the name of the weekday using label
wday(datetime, label = TRUE, abbr = FALSE)
```

##Rounding
* floor_date
* round_date
* ceiling_date
```{r}
flights_dt %>% 
  count(week = floor_date(dep_time, "week")) %>% 
  ggplot(aes(week, n)) +
    geom_line()
```

##Setting components

You can modify the component of a date time as follows
```{r}
(datetime <- ymd_hms("2016-07-08 12:34:56"))
#> [1] "2016-07-08 12:34:56 UTC"

year(datetime) <- 2020
datetime
#> [1] "2020-07-08 12:34:56 UTC"
month(datetime) <- 01
datetime
#> [1] "2020-01-08 12:34:56 UTC"
hour(datetime) <- hour(datetime) + 1
datetime
```

Setting multiple components at once
But if values are too big, they will roll over
```{r}
update(datetime, year = 2020, month = 2, mday = 2, hour = 2)

#Roll over
ymd("2015-02-01") %>% 
  update(mday = 30)
#> [1] "2015-03-02"
ymd("2015-02-01") %>% 
  update(hour = 400)
```

#Time spans

Definition

* durations, which represent an exact number of seconds.
* periods, which represent human units like weeks and months.
* intervals, which represent a starting and ending point.

##Duration

```{r}
h_age <- today() - ymd(19791014)
h_age

as.duration(h_age)
```

Creating durations using different constructors
```{r}
dseconds(15)

dminutes(10)

dhours(c(12, 24))

ddays(0:5)

dweeks(3)

dyears(1)
```

Adding duration to a date time
```{r}
dyears(1) + dweeks(12) + dhours(15)

#Adding a duration to a date time (DST occurs and hour get shifted !!)
one_pm <- ymd_hms("2016-03-12 13:00:00", tz = "America/New_York")
one_pm + ddays(1)
```

##Period 

To solve this problem, lubridate provides periods. Periods are time spans but don’t have a fixed length in seconds, instead they work with “human” times, like days and months. That allows them work in a more intuitive way:
```{r}
one_pm

one_pm + days(1)
```

Constructing components of periods
```{r}
seconds(15)

minutes(10)

hours(c(12, 24))

days(7)

months(1:6)

weeks(3)

years(1)

#Adding components
days(50) + hours(25) + minutes(2)
```

Comparison between duration and period
```{r}
# A leap year
ymd("2016-01-01") + dyears(1)

ymd("2016-01-01") + years(1)


# Daylight Savings Time
one_pm + ddays(1)

one_pm + days(1)

```

Example of recoding date time for overnight flights
```{r}
flights_dt <- flights_dt %>% 
  mutate(
    overnight = arr_time < dep_time,
    arr_time = arr_time + days(overnight * 1),
    sched_arr_time = sched_arr_time + days(overnight * 1)
  )
```

##Interval

What should years(1) / days(1) return? Well, if the year was 2015 it should return 365, but if it was 2016, it should return 366!

To overcome this issue, we can proceed as follows
```{r}
#Option 1
next_year <- today() + years(1)
(today() %--% next_year) / ddays(1)

#Option 2
(today() %--% next_year) %/% days(1)
```

#Time zones

```{r}
#Local time zone
Sys.timezone()
```

lubridate uses UTC
```{r}
(x1 <- ymd_hms("2015-06-01 12:00:00", tz = "America/New_York"))
#> [1] "2015-06-01 12:00:00 EDT"
(x2 <- ymd_hms("2015-06-01 18:00:00", tz = "Europe/Copenhagen"))
#> [1] "2015-06-01 18:00:00 CEST"
(x3 <- ymd_hms("2015-06-02 04:00:00", tz = "Pacific/Auckland"))

x1 - x2
```

Keep the same time in other timezones

```{r}
x4 <- c(x1, x2, x3)
x4a <- with_tz(x4, tzone = "Australia/Lord_Howe")
x4a
```

Force to keep the same component but in a different time zone
Thus, the resulting time corresponds to a different moment !
```{r}
x4 <- c(x1, x2, x3)
x4b <- force_tz(x4, tzone = "Australia/Lord_Howe")
x4b
```

