---
title: "Iterations"
author: "Thomas Laurent"
date: "2017年7月2日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Loading packages
library(datasets)
library(tidyverse)
```

#Iteration

One tool for reducing duplication is functions, which allows to reduce code length by allowing to perform iterations using these functions. In R, there are several ways to perform iterations as described in each section.

##For loops

Example of loop to calculate summary statistics
```{r}
df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)

#Median calculation
output <- vector("double", ncol(df))  # 1. output
for (i in seq_along(df)) {            # 2. sequence
  output[[i]] <- median(df[[i]])      # 3. body
}
output

#
```

###Modifying an existing object

Example of loop iteration used to transform original values
```{r}
df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)
rescale01 <- function(x) {
  rng <- range(x, na.rm = TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
}

#Overwriting an existing object
for (i in seq_along(df)) {
  df[[i]] <- rescale01(df[[i]])
}
```

###Unknown output length

When performing an iteration using loops, it may be better to use a list as the output rather to use a vector to avoid copying the result from previous iterations.
```{r}
means <- c(0, 1, 2)
#Not good way to do (results from the previous iteration has to be copied)
output <- double()
for (i in seq_along(means)) {
  n <- sample(100, 1)
  output <- c(output, rnorm(n, means[[i]]))
}
str(output)

#Better way to proceed
out <- vector("list", length(means))
for (i in seq_along(means)) {
  n <- sample(100, 1)
  out[[i]] <- rnorm(n, means[[i]])
}
str(out)
#> List of 3
#>  $ : num [1:83] 0.367 1.13 -0.941 0.218 1.415 ...
#>  $ : num [1:21] -0.485 -0.425 2.937 1.688 1.324 ...
#>  $ : num [1:40] 2.34 1.59 2.93 3.84 1.3 ...
str(unlist(out))
```

##While loops

```{r}
while (condition) {
  # body
}
```

##For loops vs functional

Loop can be implemented into functions to write shorter codes. However, in the code below, the codes for each statistics is very similar.
```{r}
#Tibble creation
df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)

#Function divided by statistics
col_median <- function(df) {
  output <- vector("double", length(df))
  for (i in seq_along(df)) {
    output[i] <- median(df[[i]])
  }
  output
}
col_sd <- function(df) {
  output <- vector("double", length(df))
  for (i in seq_along(df)) {
    output[i] <- sd(df[[i]])
  }
  output
}
```

To shorten the above code, another parameter "fun" is introduced in the function to determine the type of statistic calculated as an argument and not within the code of the function.
```{r}
#Shorter code with encapsulated functions
col_summary <- function(df, fun) {
  out <- vector("double", length(df))
  for (i in seq_along(df)) {
    out[i] <- fun(df[[i]])
  }
  out
}
col_summary(df, median)
#> [1]  0.237 -0.218  0.254 -0.133
col_summary(df, mean)
#> [1]  0.2026 -0.2068  0.1275 -0.0917
```

##Map functions
When looping over a vector, instead using "for" loops, map functions can be used. In this case, the type of output has to be specified as indicated below.

* map() makes a list.
* map_lgl() makes a logical vector.
* map_int() makes an integer vector.
* map_dbl() makes a double vector.
* map_chr() makes a character vector.

```{r}
map_dbl(df, mean)

map_dbl(df, median)

map_dbl(df, sd)

#Using pipelines
df %>% map_dbl(mean)

df %>% map_dbl(median)

df %>% map_dbl(sd)
```

###Shortcuts

Example of map function usage to build 
```{r}
#Simple input
models <- mtcars %>% 
  split(.$cyl) %>% 
  map(function(df) lm(mpg ~ wt, data = df))

#Same operations as above but shortened code
models <- mtcars %>% 
  split(.$cyl) %>% 
  map(~lm(mpg ~ wt, data = .))
```


###Dealing with errors

Use safely to get the error message on a particular map functions operation.
transpose function of purrr package allows to make the output clearer.
Possibly function always returns a value but if an error occurs, it will set the value as NA.
```{r}
x <- list(1, 10, "a")
y <- x %>% map(safely(log))
str(y)

y <- y %>% transpose()
str(y)

is_ok <- y$error %>% map_lgl(is_null)
x[!is_ok]

#Using possibly

x <- list(1, 10, "a")
x %>% map_dbl(possibly(log, NA_real_))
```

##Mapping over multiple arguments

Varying two arguments at the same time (ex: mean and sd)
map2 function can be used to apply a function using two parallel vectors.
pmap function performs as map2 but is used when there are more than three parameter vectors.

```{r}
#Only mean
mu <- list(5, 10, -3)
mu %>% 
  map(rnorm, n = 5) %>% 
  str()

#Both mean and sd
sigma <- list(1, 5, 10)
seq_along(mu) %>% 
  map(~rnorm(5, mu[[.]], sigma[[.]])) %>% 
  str()

##Using map2

map2(mu, sigma, rnorm, n = 5) %>% str()

#More arguments
n <- list(1, 3, 5)
args1 <- list(n, mu, sigma)
args1 %>%
  pmap(rnorm) %>% 
  str()
```

##Invoking different functions

Use different lists of parameters in combination with different functions can be performed using invoke_map function
```{r}
f <- c("runif", "rnorm", "rpois")
param <- list(
  list(min = -1, max = 1), 
  list(sd = 5), 
  list(lambda = 10)
)

invoke_map(f, param, n = 5) %>% str()
```

Perform the same operation using tribbles

```{r}
sim <- tribble(
  ~f,      ~params,
  "runif", list(min = -1, max = 1),
  "rnorm", list(sd = 5),
  "rpois", list(lambda = 10)
)
sim %>% 
  mutate(sim = invoke_map(f, params, n = 10))
```

##Walk

Use walk functions if you want to iterate for on a function side effect and not its return values
It is particularly useful for saving file or images.
```{r}
x <- list(1, "a", 3)

x %>% 
  walk(print)
```

Example when saving images of multiple plots
```{r}
library(ggplot2)
plots <- mtcars %>% 
  split(.$cyl) %>% 
  map(~ggplot(., aes(mpg, wt)) + geom_point())
paths <- stringr::str_c(names(plots), ".pdf")

pwalk(list(paths, plots), ggsave, path = tempdir())
```

##Other types of loops

* keep: keep the data if the predicate is TRUE
* discard: delete the data if the predicate is TRUE
* some: at least one of the predicate is TRUE
* every: all the predicates are TRUE
* detect: keep the observation of the first predicate which is TRUE
* detect_index: extract the index of the observation of the first predicate which is TRUE
* head_while: take elements from the start of a vector while a predicate is true
* tail_while: take elements from the end of a vector while a predicate is true:
```{r}
iris %>% 
  keep(is.factor) %>% 
  str()
#> 'data.frame':    150 obs. of  1 variable:
#>  $ Species: Factor w/ 3 levels "setosa","versicolor",..: 1 1 1 1 1 1 1 1 1 1 ...

iris %>% 
  discard(is.factor) %>% 
  str()

x <- list(1:5, letters, list(10))

x %>% 
  some(is_character)
#> [1] TRUE

x %>% 
  every(is_vector)


x <- sample(10)
x

x %>% 
  detect(~ . > 5)

x %>% 
  detect_index(~ . > 5)

x %>% 
  head_while(~ . > 5)

x %>% 
  tail_while(~ . > 5)
```

###Reduce and accumulate

Sometimes you have a complex list that you want to reduce to a simple list by repeatedly applying a function that reduces a pair to a singleton. This is useful if you want to apply a two-table dplyr verb to multiple tables.

Example: performing a full join on multiple data frames
```{r}
dfs <- list(
  age = tibble(name = "John", age = 30),
  sex = tibble(name = c("John", "Mary"), sex = c("M", "F")),
  trt = tibble(name = "Mary", treatment = "A")
)

dfs %>% reduce(full_join)
```

Accumulate is similar to reduce since it repeats an operation until there is only one element left. It keeps the intermediate results. It is particularly useful to perform cumulative sums.
```{r}
x <- sample(10)
x
x %>% accumulate(`+`)
```

