---
title: "Models"
author: "Thomas Laurent"
date: "2017年7月3日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Loading packages
library(tidyverse)
library(splines)
library(modelr)
options(na.action = na.warn)
```

#Basics

Calculating distances from the model (mean squared error)

```{r}

models <- tibble(
  a1 = runif(250, -20, 40),
  a2 = runif(250, -5, 5)
)

model1 <- function(a, data) {
  a[1] + data$x * a[2]
}
model1(c(7, 1.5), sim1)

measure_distance <- function(mod, data) {
  diff <- data$y - model1(mod, data)
  sqrt(mean(diff ^ 2))
}
measure_distance(c(7, 1.5), sim1)

sim1_dist <- function(a1, a2) {
  measure_distance(c(a1, a2), sim1)
}

models <- models %>% 
  mutate(dist = purrr::map2_dbl(a1, a2, sim1_dist))
models
```

Graphs

```{r}
ggplot(sim1, aes(x, y)) + 
  geom_point(size = 2, colour = "grey30") + 
  geom_abline(
    aes(intercept = a1, slope = a2, colour = -dist), 
    data = filter(models, rank(dist) <= 10)
  )
```

```{r}
grid <- expand.grid(
  a1 = seq(-5, 20, length = 25),
  a2 = seq(1, 3, length = 25)
  ) %>% 
  mutate(dist = purrr::map2_dbl(a1, a2, sim1_dist))
```

##Newton-Raphson method

```{r}
best <- optim(c(0, 0), measure_distance, data = sim1)
best$par

ggplot(sim1, aes(x, y)) + 
  geom_point(size = 2, colour = "grey30") + 
  geom_abline(intercept = best$par[1], slope = best$par[2])
```

##Linear models

```{r}
sim1_mod <- lm(y ~ x, data = sim1)
coef(sim1_mod)

#Adding predictions to a tibble-----------------------------------
grid <- sim1 %>% 
  data_grid(x) 
grid

grid <- grid %>% 
  add_predictions(sim1_mod) 
grid

##Visualization
ggplot(sim1, aes(x)) +
  geom_point(aes(y = y)) +
  geom_line(aes(y = pred), data = grid, colour = "red", size = 1)

#Adding residuals-------------------------------------------------

sim1 <- sim1 %>% 
  add_residuals(sim1_mod)

ggplot(sim1, aes(x, resid)) + 
  geom_ref_line(h = 0) +
  geom_point() 
```

##Formulas and model families

Observe formula
```{r}
df <- tribble(
  ~y, ~x1, ~x2,
  4, 2, 5,
  5, 1, 6
)
model_matrix(df, y ~ x1)

#Without intercept
model_matrix(df, y ~ x1 - 1)

#Two variables
model_matrix(df, y ~ x1 + x2)

#Factor
df <- tribble(
  ~ sex, ~ response,
  "male", 1,
  "female", 2,
  "male", 1
)
model_matrix(df, response ~ sex)
```


###Interactions

```{r}
mod1 <- lm(y ~ x1 + x2, data = sim3)
mod2 <- lm(y ~ x1 * x2, data = sim3)

#Output

grid <- sim3 %>% 
  data_grid(x1, x2) %>% 
  spread_predictions(mod1, mod2)
grid

#Graphics

sim3 <- sim3 %>% 
  gather_residuals(mod1, mod2)

ggplot(sim3, aes(x1, resid, colour = x2)) + 
  geom_point() + 
  facet_grid(model ~ x2)
```

###Interactions (two continuous variables)

```{r}
mod1 <- lm(y ~ x1 + x2, data = sim4)
mod2 <- lm(y ~ x1 * x2, data = sim4)

grid <- sim4 %>% 
  data_grid(
    x1 = seq_range(x1, 5), 
    x2 = seq_range(x2, 5) 
  ) %>% 
  gather_predictions(mod1, mod2)
grid

#Usaeg of seq_range to generate a grid between min and max
seq_range(c(0.0123, 0.923423), n = 5)

seq_range(c(0.0123, 0.923423), n = 5, pretty = TRUE)
```

###Transformations

```{r}
#Use of I for transformations
df <- tribble(
  ~y, ~x,
   1,  1,
   2,  2, 
   3,  3
)

model_matrix(df, y ~ I(x^2) + x)

#Polynomial

model_matrix(df, y ~ poly(x, 2))
```

Approximation by splines
```{r}
sim5 <- tibble(
  x = seq(0, 3.5 * pi, length = 50),
  y = 4 * sin(x) + rnorm(length(x))
)

mod1 <- lm(y ~ ns(x, 1), data = sim5)
mod2 <- lm(y ~ ns(x, 2), data = sim5)
mod3 <- lm(y ~ ns(x, 3), data = sim5)
mod4 <- lm(y ~ ns(x, 4), data = sim5)
mod5 <- lm(y ~ ns(x, 5), data = sim5)

grid <- sim5 %>% 
  data_grid(x = seq_range(x, n = 50, expand = 0.1)) %>% 
  gather_predictions(mod1, mod2, mod3, mod4, mod5, .pred = "y")

ggplot(sim5, aes(x, y)) + 
  geom_point() +
  geom_line(data = grid, colour = "red") +
  facet_wrap(~ model)
```

##Missing values

```{r}
df <- tribble(
  ~x, ~y,
  1, 2.2,
  2, NA,
  3, 3.5,
  4, 8.3,
  NA, 10
)

#Suppress warnings of missing values
mod <- lm(y ~ x, data = df, na.action = na.exclude)

#Number of observations used in the model
nobs(mod)
```

#Building models

```{r}
#Packages
library(tidyverse)
library(modelr)
options(na.action = na.warn)

library(nycflights13)
library(lubridate)
```

```{r}
#Graphs
ggplot(diamonds, aes(cut, price)) + geom_boxplot()
ggplot(diamonds, aes(color, price)) + geom_boxplot()
ggplot(diamonds, aes(clarity, price)) + geom_boxplot()
```

Focus on partial population

```{r}
diamonds2 <- diamonds %>% 
  filter(carat <= 2.5) %>% 
  mutate(lprice = log2(price), lcarat = log2(carat))

#Plotting bins
ggplot(diamonds2, aes(lcarat, lprice)) + 
  geom_hex(bins = 50)
```

##Modeling

```{r}
mod_diamond <- lm(lprice ~ lcarat, data = diamonds2)

grid <- diamonds2 %>% 
  data_grid(carat = seq_range(carat, 20)) %>% 
  mutate(lcarat = log2(carat)) %>% 
  add_predictions(mod_diamond, "lprice") %>% 
  mutate(price = 2 ^ lprice)

ggplot(diamonds2, aes(carat, price)) + 
  geom_hex(bins = 50) + 
  geom_line(data = grid, colour = "red", size = 1)
```

Visualizing the residuals

```{r}
diamonds2 <- diamonds2 %>% 
  add_residuals(mod_diamond, "lresid")

ggplot(diamonds2, aes(lcarat, lresid)) + 
  geom_hex(bins = 50)

#Boxplots
ggplot(diamonds2, aes(cut, lresid)) + geom_boxplot()
ggplot(diamonds2, aes(color, lresid)) + geom_boxplot()
ggplot(diamonds2, aes(clarity, lresid)) + geom_boxplot()
```

#Many models

```{r}
#Accessing data of progression of countries over time
library(gapminder)
gapminder
```

```{r}
gapminder %>% 
  ggplot(aes(year, lifeExp, group = country)) +
    geom_line(alpha = 1/3)
```

##Performing a linear regression using NZ data

```{r}
nz <- filter(gapminder, country == "New Zealand")
nz %>% 
  ggplot(aes(year, lifeExp)) + 
  geom_line() + 
  ggtitle("Full data = ")

nz_mod <- lm(lifeExp ~ year, data = nz)
nz %>% 
  add_predictions(nz_mod) %>%
  ggplot(aes(year, pred)) + 
  geom_line() + 
  ggtitle("Linear trend + ")

nz %>% 
  add_residuals(nz_mod) %>% 
  ggplot(aes(year, resid)) + 
  geom_hline(yintercept = 0, colour = "white", size = 3) + 
  geom_line() + 
  ggtitle("Remaining pattern")
```

##Nested data

Use nest() to create a nested data frame
```{r}
by_country <- gapminder %>% 
  group_by(country, continent) %>% 
  nest()

#Observe the contents of the nested data frame
by_country$data[[1]]
```

##List-columns

Using list-columns, iterative modeling can be performed using map() function.
```{r}
country_model <- function(df) {
  lm(lifeExp ~ year, data = df)
}

models <- map(by_country$data, country_model)

by_country <- by_country %>% 
  mutate(model = map(data, country_model))
```

##Unnesting

Adding the residuals

```{r}
by_country <- by_country %>% 
  mutate(
    resids = map2(data, model, add_residuals)
  )
by_country
```

```{r}
resids <- unnest(by_country, resids)
resids
```

Plotting lines by country

```{r}
resids %>% 
  ggplot(aes(year, resid)) +
    geom_line(aes(group = country), alpha = 1 / 3) + 
    geom_smooth(se = FALSE)
```

```{r}
resids %>% 
  ggplot(aes(year, resid, group = country)) +
    geom_line(alpha = 1 / 3) + 
    facet_wrap(~continent)
```

##Model quality

Turn models into tidy data
```{r}
broom::glance(nz_mod)
```

```{r}
glance <- by_country %>% 
  mutate(glance = map(model, broom::glance)) %>% 
  unnest(glance, .drop = TRUE)
glance
```

Plotting R2 by continent
```{r}
glance %>% 
  ggplot(aes(continent, r.squared)) + 
    geom_jitter(width = 0.5)
```

Plotting badly fitted countries
```{r}
bad_fit <- filter(glance, r.squared < 0.25)

gapminder %>% 
  semi_join(bad_fit, by = "country") %>% 
  ggplot(aes(year, lifeExp, colour = country)) +
    geom_line()
```

##List columns

List columns can be created using tribble or tibble

```{r}
tribble(
   ~x, ~y,
  1:3, "1, 2",
  3:5, "3, 4, 5"
)

tibble(
  x = list(1:3, 3:5), 
  y = c("1, 2", "3, 4, 5")
)
```

###Principle to create a list-column

*With tidyr::nest() to convert a grouped data frame into a nested data frame where you have list-column of data frames.

*With mutate() and vectorised functions that return a list.

*With summarise() and summary functions that return multiple results.

###Using nest()

```{r}
gapminder %>% 
  group_by(country, continent) %>% 
  nest()
```

```{r}
gapminder %>% 
  nest(year:gdpPercap)
```

###From vectorized function

```{r}
df <- tribble(
  ~x1,
  "a,b,c", 
  "d,e,f,g"
) 

df %>% 
  mutate(x2 = stringr::str_split(x1, ","))

#Handle with unnest()
df %>% 
  mutate(x2 = stringr::str_split(x1, ",")) %>% 
  unnest()
```

```{r}
sim <- tribble(
  ~f,      ~params,
  "runif", list(min = -1, max = -1),
  "rnorm", list(sd = 5),
  "rpois", list(lambda = 10)
)

sim %>%
  mutate(sims = invoke_map(f, params, n = 10))
```

###From multivalued summaries

list() allows to store multivalues in the tibble
```{r}
mtcars %>% 
  group_by(cyl) %>% 
  summarise(q = list(quantile(mpg)))
```

```{r}
probs <- c(0.01, 0.25, 0.5, 0.75, 0.99)
mtcars %>% 
  group_by(cyl) %>% 
  summarise(p = list(probs), q = list(quantile(mpg, probs))) %>% unnest()
```

###From a name list

```{r}
x <- list(
  a = 1:5,
  b = 3:4, 
  c = 5:6
) 

df <- enframe(x)
```

```{r}
df %>% 
  mutate(
    smry = map2_chr(name, value, ~ stringr::str_c(.x, ": ", .y[1]))
  )
```

##Simplifying list-columns

map_* functions can be used to return a single value.

```{r}
df <- tribble(
  ~x,
  letters[1:5],
  1:3,
  runif(5)
)
  
df %>% mutate(
  type = map_chr(x, typeof),
  length = map_int(x, length)
)
```

```{r}
df <- tribble(
  ~x,
  list(a = 1, b = 2),
  list(a = 2, c = 4)
)
df %>% mutate(
  a = map_dbl(x, "a"),
  b = map_dbl(x, "b", .null = NA_real_)
)
```

###Unnesting

```{r}
tibble(x = 1:2, y = list(1:4, 1))
```

##Making tidy data with broom

* broom::glance(model) returns a row for each model. Each column gives a model summary: either a measure of model quality, or complexity, or a combination of the two.

* broom::tidy(model) returns a row for each coefficient in the model. Each column gives information about the estimate or its variability.

* broom::augment(model, data) returns a row for each row in data, adding extra values like residuals, and influence statistics.

```{r}
glance <- by_country %>% 
  mutate(glance = map(model, broom::tidy))

glance %>% select(country,continent,resids) %>% unnest()
```

