---
title: "Programmation"
author: "Thomas Laurent"
date: "2017年7月2日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Loading packages
library(datasets)
library(tidyverse)
```

#Functions

##Switch function

One useful technique is the switch() function. It allows you to evaluate selected code based on position or name.

```{r}
operate=function(x, y, op) {
  switch(op,
     plus = x + y,
     minus = x - y,
     times = x * y,
     divide = x / y,
     stop("Unknown op!")
  ) }
operate(3,2,"plus")
```

##Writing pipeline functions

```{r}
show_missings <- function(df) {
  n <- sum(is.na(df))
  cat("Missing values: ", n, "\n", sep = "")
 #Invisible to not return the input 
  invisible(df)
}

#dummy data frame
data=data.frame(
  a=rnorm(10)*100,
  b=rnorm(10)*100
)
data[1,2]=NA

#Executing the function
show_missings(data)
```

Combining in a pipe

```{r}
mtcars %>% 
  show_missings() %>% 
  mutate(mpg = ifelse(mpg < 20, NA, mpg)) %>% 
  show_missings() 
```

#Vectors

```{r}
#Determining the type

typeof(letters)
typeof(1)
typeof(1L)

#
x <- "This is a reasonably long string."
pryr::object_size(x)
```

##Naming vectors

```{r}
set_names(1:3, c("a", "b", "c"))
```

##Lists

```{r}
#Creating list
x <- list(1, 2, 3)
str(x)

#Combined list
z <- list(list(1, 2), list(3, 4))
str(z)
```

Subsetting

```{r}
a <- list(a = 1:3, b = "a string", c = pi, d = list(-1, -5))

#[] returns a list
str(a[1:2])

#returns a object without hierarchy
str(a[[1]])

a$a

a[["a"]]
```

##Attributes

```{r}
attr(x, "greeting")
#> NULL
attr(x, "greeting") <- "Hi!"
attr(x, "farewell") <- "Bye!"
attributes(x)
```

Case of tibbles

```{r}
tb <- tibble::tibble(x = 1:5, y = 5:1)
typeof(tb)

attributes(tb)

```

#Iteration

##For loops

```{r}
df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)

#Median calculation
output <- vector("double", ncol(df))  # 1. output
for (i in seq_along(df)) {            # 2. sequence
  output[[i]] <- median(df[[i]])      # 3. body
}
output

#
```

###Modifying an existing object

```{r}
df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)
rescale01 <- function(x) {
  rng <- range(x, na.rm = TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
}

#Overwriting an existing object
for (i in seq_along(df)) {
  df[[i]] <- rescale01(df[[i]])
}
```

###Unknown output length

```{r}
means <- c(0, 1, 2)
#Not good way to do (results from the previous iteration has to be copied)
output <- double()
for (i in seq_along(means)) {
  n <- sample(100, 1)
  output <- c(output, rnorm(n, means[[i]]))
}
str(output)

#Better way to proceed
out <- vector("list", length(means))
for (i in seq_along(means)) {
  n <- sample(100, 1)
  out[[i]] <- rnorm(n, means[[i]])
}
str(out)
#> List of 3
#>  $ : num [1:83] 0.367 1.13 -0.941 0.218 1.415 ...
#>  $ : num [1:21] -0.485 -0.425 2.937 1.688 1.324 ...
#>  $ : num [1:40] 2.34 1.59 2.93 3.84 1.3 ...
str(unlist(out))
```

##While loops

```{r}
while (condition) {
  # body
}
```

##For loops vs functional

```{r}
#Tibble creation
df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)

#Function divided by statistics
col_median <- function(df) {
  output <- vector("double", length(df))
  for (i in seq_along(df)) {
    output[i] <- median(df[[i]])
  }
  output
}
col_sd <- function(df) {
  output <- vector("double", length(df))
  for (i in seq_along(df)) {
    output[i] <- sd(df[[i]])
  }
  output
}
```

```{r}
#Shorter code with encapsulated functions
col_summary <- function(df, fun) {
  out <- vector("double", length(df))
  for (i in seq_along(df)) {
    out[i] <- fun(df[[i]])
  }
  out
}
col_summary(df, median)
#> [1]  0.237 -0.218  0.254 -0.133
col_summary(df, mean)
#> [1]  0.2026 -0.2068  0.1275 -0.0917
```

##Map functions

* map() makes a list.
* map_lgl() makes a logical vector.
* map_int() makes an integer vector.
* map_dbl() makes a double vector.
* map_chr() makes a character vector.

```{r}
map_dbl(df, mean)

map_dbl(df, median)

map_dbl(df, sd)

#Using pipelines
df %>% map_dbl(mean)

df %>% map_dbl(median)

df %>% map_dbl(sd)
```

###Shortcuts

```{r}
#Simple input
models <- mtcars %>% 
  split(.$cyl) %>% 
  map(function(df) lm(mpg ~ wt, data = df))

#Shortened input
models <- mtcars %>% 
  split(.$cyl) %>% 
  map(~lm(mpg ~ wt, data = .))
```


###Dealing with errors

Use safely to get the error message
```{r}
x <- list(1, 10, "a")
y <- x %>% map(safely(log))
str(y)

y <- y %>% transpose()
str(y)

is_ok <- y$error %>% map_lgl(is_null)
x[!is_ok]

#Using possibly

x <- list(1, 10, "a")
x %>% map_dbl(possibly(log, NA_real_))
```

##Mapping over multiple arguments

Varying two arguments at the same time (ex: mean and sd)

```{r}
#Only mean
mu <- list(5, 10, -3)
mu %>% 
  map(rnorm, n = 5) %>% 
  str()

#Both mean and sd
sigma <- list(1, 5, 10)
seq_along(mu) %>% 
  map(~rnorm(5, mu[[.]], sigma[[.]])) %>% 
  str()

##Using map2

map2(mu, sigma, rnorm, n = 5) %>% str()

#More arguments
n <- list(1, 3, 5)
args1 <- list(n, mu, sigma)
args1 %>%
  pmap(rnorm) %>% 
  str()
```

##Invoking different functions

Use different lists of parameters in combination with different functions
```{r}
f <- c("runif", "rnorm", "rpois")
param <- list(
  list(min = -1, max = 1), 
  list(sd = 5), 
  list(lambda = 10)
)

invoke_map(f, param, n = 5) %>% str()
```

Perform the same operation using tribbles

```{r}
sim <- tribble(
  ~f,      ~params,
  "runif", list(min = -1, max = 1),
  "rnorm", list(sd = 5),
  "rpois", list(lambda = 10)
)
sim %>% 
  mutate(sim = invoke_map(f, params, n = 10))
```

##Walk

```{r}
x <- list(1, "a", 3)

x %>% 
  walk(print)
```


```{r}
library(ggplot2)
plots <- mtcars %>% 
  split(.$cyl) %>% 
  map(~ggplot(., aes(mpg, wt)) + geom_point())
paths <- stringr::str_c(names(plots), ".pdf")

pwalk(list(paths, plots), ggsave, path = tempdir())
```

##Other types of loops

* keep: keep the data if the predicate is TRUE
* discard: delete the data if the predicate is TRUE
* some: at least one of the predicate is TRUE
* every: all the predicates are TRUE
* detect: keep the observation of the first predicate which is TRUE
* detect_index: extract the index of the observation of the first predicate which is TRUE
* head_while: take elements from the start of a vector while a predicate is true
* tail_while: take elements from the end of a vector while a predicate is true:
```{r}
iris %>% 
  keep(is.factor) %>% 
  str()
#> 'data.frame':    150 obs. of  1 variable:
#>  $ Species: Factor w/ 3 levels "setosa","versicolor",..: 1 1 1 1 1 1 1 1 1 1 ...

iris %>% 
  discard(is.factor) %>% 
  str()

x <- list(1:5, letters, list(10))

x %>% 
  some(is_character)
#> [1] TRUE

x %>% 
  every(is_vector)


x <- sample(10)
x


x %>% 
  detect(~ . > 5)
#> [1] 8

x %>% 
  detect_index(~ . > 5)
#> [1] 1

x %>% 
  head_while(~ . > 5)
#> [1] 8 7

x %>% 
  tail_while(~ . > 5)
```

###Reduce and accumulate

Sometimes you have a complex list that you want to reduce to a simple list by repeatedly applying a function that reduces a pair to a singleton.

Example: performing a full join on multiple data frames
```{r}
dfs <- list(
  age = tibble(name = "John", age = 30),
  sex = tibble(name = c("John", "Mary"), sex = c("M", "F")),
  trt = tibble(name = "Mary", treatment = "A")
)

dfs %>% reduce(full_join)
```

Accumulate is similar to reduce but keeps the intermediate results
```{r}
x <- sample(10)
x
x %>% accumulate(`+`)
```

