---
title: "Relational data"
author: "Thomas Laurent"
date: "2017年7月1日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Loading package
library(tidyverse)
library(nycflights13)
```

There exist two types of key for tables in relational database:

* A primary key uniquely identifies an observation in its own table. For example, planes$tailnum is a primary key because it uniquely identifies each plane in the planes table.

* A foreign key uniquely identifies an observation in another table. For example, the flights$tailnum is a foreign key because it appears in the flights table where it matches each flight to a unique plane.

#Checking primary keys

The unicity of primary key can be checked as follows.
```{r}
#Primary key
planes %>% 
  count(tailnum) %>% 
  filter(n > 1)
```

#Mutating joins

Table join can be done using "X"_join() functions included in dplyr.
Mutating joins allows to add an additional variable using an external table using match() functions by matching the key between the original table and the external table.
```{r}
#using dplyr
flights %>%
  select(-origin, -dest) %>% 
  left_join(airlines, by = "carrier")

#using mutate
flights %>%
  select(-origin, -dest) %>% 
  mutate(name = airlines$name[match(carrier, airlines$carrier)])
```

##Inner join

```{r}
x <- tribble(
  ~key, ~val_x,
     1, "x1",
     2, "x2",
     3, "x3"
)
y <- tribble(
  ~key, ~val_y,
     1, "y1",
     2, "y2",
     4, "y3"
)
```


```{r}
x %>% 
  inner_join(y, by = "key")
```

##Left join

```{r}
left_join(x, y, by = "key")
```

##Duplicated keys

Issues can happen if the key in each table is not unique as below. Thus, you should be careful about this point when joining.
```{r}
x <- tribble(
  ~key, ~val_x,
     1, "x1",
     2, "x2",
     2, "x3",
     1, "x4"
)
y <- tribble(
  ~key, ~val_y,
     1, "y1",
     2, "y2"
)
left_join(x, y, by = "key")
```

##Options

* by: use all common variables if NULL

```{r}
#key argument when names differ
flights %>% 
  left_join(airports, by=c("dest" = "faa"))
```

##Semi-join
The principle of semi-joins is as follows:
* semi_join(x, y) keeps all observations in x that have a match in y.
* anti_join(x, y) drops all observations in x that have a match in y.

```{r}
top_dest <- flights %>%
  count(dest, sort = TRUE) %>%
  head(10)

#Semi-join using filter
flights %>% 
  filter(dest %in% top_dest$dest)
```

##Anti-join

Select observations that does not meet the criteria

```{r}
#Anti-join
flights %>%
  anti_join(planes, by = "tailnum") %>%
  count(tailnum, sort = TRUE)
```

#Set operations

Union, intersection and differences in observations between dataframe with the same structure
```{r}
df1 <- tribble(
  ~x, ~y,
   1,  1,
   2,  1
)
df2 <- tribble(
  ~x, ~y,
   1,  1,
   1,  2
)

#Intersection
intersect(df1, df2)

#Union
union(df1,df2)

#Difference
##Returns the observation in df1 (first declared data.frame) that are not in df2
setdiff(df1,df2)
```

#Other implementations

##Comparison dplyr and merge
dplyr functions "X"_join are equivalent to merge function in R base. However, dplyr joins are much more faster than their respective equivalents using merge function.

* inner_join(x, y)	equivalent to merge(x, y)
* left_join(x, y)	equivalent to merge(x, y, all.x = TRUE)
* right_join(x, y) equivalent to merge(x, y, all.y = TRUE),
* full_join(x, y)	equivalent merge(x, y, all.x = TRUE, all.y = TRUE)

##Filtering joins
Affect only the observations not the data