---
title: "Rmarkdown"
output: html_document
params:
  my_class: "suv"
---

<style type="text/css">

body{ /* Normal  */
      font-size: 12px;
  }
td {  /* Table  */
  font-size: 8px;
}
h1.title {
  font-size: 38px;
  color: DarkRed;
}
h1 { /* Header 1 */
  font-size: 28px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 22px;
  color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>

#Example
```{r setup, include = FALSE}
library(ggplot2)
library(dplyr)
library(magrittr)

smaller <- diamonds %>% 
  filter(carat <= 2.5)
```

We have data about `r nrow(diamonds)` diamonds. Only 
`r nrow(diamonds) - nrow(smaller)` are larger than
2.5 carats. The distribution of the remainder is shown
below:

```{r, echo = FALSE}
smaller %>% 
  ggplot(aes(carat)) + 
  geom_freqpoly(binwidth = 0.01)
```

#Structure

* YAML header : the structure surrounded by ---
  ---
  ---

* Chunks : the structure surrounded by ```

* Text

#Code execution

Ctrl+shit+Enter on RStudio in Windows

#Formatting

Text formatting 
------------------------------------------------------------

*italic*  or _italic_
**bold**   __bold__
`code`
superscript^2^ and subscript~2~

Headings
------------------------------------------------------------

# 1st Level Header

## 2nd Level Header

### 3rd Level Header

Lists
------------------------------------------------------------

*   Bulleted list item 1

*   Item 2

    * Item 2a

    * Item 2b

1.  Numbered list item 1

1.  Item 2. The numbers are incremented automatically in the output.

Links and images
------------------------------------------------------------

<http://example.com>



Tables 
------------------------------------------------------------

First Header  | Second Header
------------- | -------------
Content Cell  | Content Cell
Content Cell  | Content Cell

#Code chunk

* Adding a code chunk : Ctrl+alt+I

## Code chunk options

* eval=FALSE : does not run the code
* include=FALSE : run the code but does not show the code and the results
* echo=FALSE : prevents the code to appear in the output document
* message=FALSE : does not show messages
* warning=FALSE : does not show warnings
* results='hide' : hide results
* fig.show='hide' : hide plots
* cache = TRUE : store the results of the chunk and rerun the chunk in the next execution only if the code has been changed.

## Table display

Use of kable() function from knitr package allows to obtain nicer outputs.
```{r}
knitr::kable(
  mtcars[1:5, ], 
  caption = "A knitr kable."
)
```

##Global options

To st global options for chunk, use opts_cunk$set() function as shown in the example.

```{r}
knitr::opts_chunk$set(
  echo = FALSE
)
```

#Inline code

You can use mark to display computed numbers directly in the text. Like this: `r nrow(diamonds)`

##Formats
 If inserting numbers in the text, format can be changed as follows for large numbers.
```{r}
comma <- function(x) format(x, digits = 2, big.mark = ",")
comma(3452345)
comma(.12358124331)
```


#YAML header

Parameters can be set directly in the Rmarkdown header.

For example in this document, header was set as : my_class: "suv" .

```{r}
library(ggplot2)
library(dplyr)
#The parameter my_class is used in the following statement
class <- mpg %>% filter(class == params$my_class)

ggplot(class, aes(displ, hwy)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

R code can be executed in the header using !r operator.

params:
  start: !r lubridate::ymd("2015-01-01")
  snapshot: !r lubridate::ymd_hms("2015-01-01 12:30:00")
  
This following code can be read using hte following code.
pwalk() allows for multiple input simultaneously.
```{r,eval=FALSE}
reports <- tibble(
  class = unique(mpg$class),
  filename = stringr::str_c("fuel-economy-", class, ".html"),
  params = purrr::map(class, ~ list(my_class = .))
)

reports %>% 
  select(output_file = filename, params) %>% 
  purrr::pwalk(rmarkdown::render, input = "fuel-economy.Rmd")
```

#Graphics

##Titles

Titles and footnotes can be added using labs() function.
```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  labs(
    #Adding main title
    title = "Fuel efficiency generally decreases with engine size",
    #Adding sub title
    subtitle = "Two seaters (sports cars) are an exception because of their light weight",
    #Adding footnote
    caption = "Data from fueleconomy.gov"
  )
```

Axis labels can be modified using labs() function.
```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_smooth(se = FALSE) +
  labs(
    #x-axis label
    x = "Engine displacement (L)",
    #y-axis label
    y = "Highway fuel economy (mpg)",
    #Color legend label
    colour = "Car type"
  )
```

##Annotations

Annotations can be added to the graph using geom_text().
```{r}
best_in_class <- mpg %>%
  group_by(class) %>%
  filter(row_number(desc(hwy)) == 1)

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  #Adding annotation to the graph
  geom_text(aes(label = model), data = best_in_class)
```

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_point(size = 3, shape = 1, data = best_in_class) +
  geom_label_repel(aes(label = model), data = best_in_class)
```

