---
title: "Tibbles"
output: 
  html_document:
    number_sections: true
    toc: true
---

<style type="text/css">

body{ /* Normal  */
      font-size: 12px;
  }
td {  /* Table  */
  font-size: 8px;
}
h1.title {
  font-size: 38px;
  color: DarkRed;
}
h1 { /* Header 1 */
  font-size: 28px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 22px;
  color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Loading packages
library(tidyverse)
```

#Creating tibbles

* as_tibble(): conversion of a data frame to tibble
* tibble(): creating a data frame from vectors
* tribble(): creating a data frame (column headers start with ~)

In tibbles, the type of data is not changed during the process.
```{r}
#Converting a data frame to iris
as_tibble(iris)

#Creating a tibble by combining vectors
tibble(
  x = 1:5, 
  y = 1, 
  z = x ^ 2 + y
)

#Creating a tibble by inputing each value
tribble(
  #column headers
  ~x, ~y, ~z,
  #--|--|----
  #Data input
  "a", 2, 3.6,
  "b", 1, 8.5
)
```

#Printing tibbles

The main difference in the usage of data frame and tibbles concerns printing and subsetting.

```{r}
  #Options of tibble printing
  #Print only n rows if more than m rows in the data
  options(tibble.print_max = n, tibble.print_min = m)
  
  #To show all rows
  options(dplyr.print_min = Inf)
```

#Subsetting

There exist several ways to extract variables from tibble.
```{r}
#Creating a test tibble
df <- tibble(
  x = runif(5),
  y = rnorm(5)
)

# Extract by name
df$x

df[["x"]]

# Extract by position (extract the first variable)
df[[1]]

```

