---
title: "Tidy data"
author: "Thomas Laurent"
date: "2017年6月29日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Loading package
library(tidyverse)
```

#Gathering

To gather columns into a new pair of variables
```{r}
#Gathering by year
table4a %>% 
  gather(`1999`, `2000`, key = "year", value = "cases")
```

Performing two gathering steps and merging datasets
```{r}
tidy4a <- table4a %>% 
  gather(`1999`, `2000`, key = "year", value = "cases")
tidy4b <- table4b %>% 
  gather(`1999`, `2000`, key = "year", value = "population")
left_join(tidy4a, tidy4b)
```

#Spreading

To put observations that are on multiple rows into one row.

```{r}
#Key :equivalent to by
#Value: equivalent to var
spread(table2, key = type, value = count)
```

#Separating one column

Separate the data from one column into two columns based on a non alphanumeric character.

```{r}
#Separate
table3 %>% 
  separate(rate, into = c("cases", "population"))

#Setting the separator

table3 %>% 
  separate(rate, into = c("cases", "population"), sep = "/",convert=TRUE)
```

```{r}
#Using position
table3 %>% 
  separate(year, into = c("century", "year"), sep = 2)
```

#Unite

It performs the opposite operation.

```{r}
#Unite
table5 %>% 
  unite(new, century, year)
```

#Missing values

* Explicitly, i.e. flagged with NA.
* Implicitly, i.e. simply not present in the data.

```{r}
stocks <- tibble(
  year   = c(2015, 2015, 2015, 2015, 2016, 2016, 2016),
  qtr    = c(   1,    2,    3,    4,    2,    3,    4),
  return = c(1.88, 0.59, 0.35,   NA, 0.92, 0.17, 2.66)
)

stocks
```


Make implicit missing values explicit

```{r}
#Implicit to explicit
stocks %>% 
  spread(year, return)

stocks %>% 
  complete(year, qtr)
```

```{r}
#Removing missing values
stocks %>% 
  spread(year, return) %>% 
  gather(year, return, `2015`:`2016`, na.rm = TRUE)

stocks %>% na.omit
```

Carry forward non missing values to fill the dataset
```{r}
treatment <- tribble(
  ~ person,           ~ treatment, ~response,
  "Derrick Whitmore", 1,           7,
  NA,                 2,           10,
  NA,                 3,           9,
  "Katherine Burke",  1,           4
)

treatment %>% fill(person)
```

#Example

```{r}
who1 <- who %>% 
  gather(new_sp_m014:newrel_f65, key = "key", value = "cases", na.rm = TRUE)

#Count the number of keys
who1 %>% 
  count(key)

#Separating relapse type and sex
who2 <- who1 %>% 
  mutate(key = stringr::str_replace(key, "newrel", "new_rel"))

who3 <- who2 %>% 
  separate(key, c("new", "type", "sexage"), sep = "_")

who4 <- who3 %>% 
  select(-new, -iso2, -iso3)

#Creating age and sex variable by splitting sexage variable
who5 <- who4 %>% 
  separate(sexage, c("sex", "age"), sep = 1)
who5
```

Performs all operation as above using pipes

```{r}
#Using pipes
who %>%
  gather(code, value, new_sp_m014:newrel_f65, na.rm = TRUE) %>% 
  mutate(code = stringr::str_replace(code, "newrel", "new_rel")) %>%
  separate(code, c("new", "var", "sexage")) %>% 
  select(-new, -iso2, -iso3) %>% 
  separate(sexage, c("sex", "age"), sep = 1)
```

